%GCE of WPI
%by Jiamin JIAN

\documentclass[12pt,a4paper]{ctexart}
\usepackage{CJK}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{float}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{amstext}
\usepackage{blkarray}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage{listings}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}

\begin{document}


\begin{center}
\textbf{Homework 4, MA573}
\vspace{8pt}

\textbf{Monotonicity in volatility}

Jiamin JIAN
\end{center}

\vspace{12pt}

$\textbf{Exercise1:}$

Prove or disprove: Suppose $f$ is convex and $X$ is submartingale, prove that  $g(t) = \mathbb E[f(X_t)]$ is increasing.

\vspace{8pt}
$\textbf{Solution:}$

This proposition is not right. We can give a counter example. Assume $f(x) = -x$, then $f(x)$ is a convex function since $\forall x_{1}, x_{2} \in \mathbb{R}$, and $\forall t \in [0, 1]$, we have $f(t x_{1} + (1-t)x_{2}) \leq t f(x_{1}) + (1-t)f(x_{2})$ holds. Then we can verify the proposotion that $g(t)$ is increasing. Since $\forall s \leq t$,
\begin{equation*}
\mathbb{E} [f(X_{t}) - f(X_{s}) | \mathcal{F}_{s}] = \mathbb{E}[-X_{t} + X_{s} | \mathcal{F}_{s}] = X_{s} - \mathbb{E}[X_{t} | \mathcal{F}_{s}].
\end{equation*}
Since $X_{t}$ is a submartigale with respect to the filtration $\{\mathcal{F}_{t}\}_{t \in \mathbb{T}}$, then we have
\begin{equation*}
\mathbb{E}[X_{t} | \mathcal{F}_{s}] \geq X_{s},
\end{equation*}
thus we know that $\forall s \leq t$, $\mathbb{E} [f(X_{t}) - f(X_{s}) | \mathcal{F}_{s}] \leq 0$. Then we take expectation on the both side and by the definition of conditional expectation, we have
$$\mathbb{E} [f(X_{t}) - f(X_{s})] \leq 0.$$
Then we know that $\forall s \leq t$, $g(t) \leq g(s)$, so that we have $g(t)$ is decreasing.

If we know $f$ is a increasing function, then by the Jensen's inequality, we have for all $s \leq t$,
$$\mathbb{E} [f(X_{t}) | \mathcal{F}_{s}] \geq f(\mathbb{E} [X_{t} | \mathcal{F}_{s}]),$$
and since $X_{t}$ is a submartingale, we have $\mathbb{E} [X_{t} | \mathcal{F}_{s}] \geq X_{s}$. As $f$ is a increasing function, then
$$\mathbb{E} [f(X_{t}) | \mathcal{F}_{s}] \geq f(\mathbb{E} [X_{t} | \mathcal{F}_{s}]) \geq f(X_{s}),$$
by taking expectation on the both side, we have
$$\mathbb{E} [f(X_{t})] \geq \mathbb{E} [f(X_{s})], \forall s \leq t.$$
Then we know $g(t)$ is increasing.

\vspace{8pt}

$\textbf{Exercise2:}$

Let $t \mapsto e^{-rt}S_t$ be a martingale, then prove that 
$$C(t) = \mathbb E[e^{-rt}(S_t - K)^+]$$ 
is increasing.

\vspace{8pt}
$\textbf{Solution:}$

Since $e^{-rt} S_{t}$ is a martingale with respect to the filtration $\{\mathcal{F}_{t}\}_{t \in \mathbb{T}}$, and the function $f(x) = x^{+}$ is a convex function, then by the Jensen's inequality in the conditional expectation, we have $\forall s \leq t$,
$$\mathbb{E} [e^{-rt}(S_{t} - K)^{+} | \mathcal{F}_{s}] \geq (\mathbb{E} [e^{-rt}(S_{t} - K) | \mathcal{F}_{s}])^{+} = (e^{-rs}S_{s} - e^{-rt} K)^{+},$$
And we have 
$$(e^{-rs}S_{s} - e^{-rt} K)^{+} = \big((e^{-rs}(S_{s} - K) + K (e^{-rs} - e^{-rt})\big)^{+} \geq (e^{-rs}(S_{s} - K))^{+}$$
since $K (e^{-rs} -  e^{-rt}) \geq 0$, $\forall s \leq t$ and $K > 0$. Thus we know that $\mathbb{E} [e^{-rt}(S_{t} - K)^{+} | \mathcal{F}_{s}] \geq (e^{-rs}(S_{s} - K))^{+}$, then we take expectation on the both side, we have that
$$\mathbb{E} [e^{-rt}(S_{t} - K)^{+}] \geq \mathbb{E} [(e^{-rs}(S_{s} - K))^{+}].$$
So $\forall s \leq t$, we have $C(s) \leq C(t)$, then $C(t)$ is increasing.

\vspace{8pt}

$\textbf{Exercise3:}$

Suppose $r = 0$ and $S$ is martingale, prove that $P(t) = \mathbb E [(S_t - K)^-]$ is increasing.

\vspace{8pt}
$\textbf{Solution:}$

Since $ S_{t}$ is a martingale with respect to the filtration $\{\mathcal{F}_{t}\}_{t \in \mathbb{T}}$, and the function $f(x) = x^{-}$ is a convex function, we have $\forall s \leq t$,
$$\mathbb{E} [(S_{t} - K)^{-} | \mathcal{F}_{s}] \geq (\mathbb{E} [S_{t} - K| \mathcal{F}_{s}])^{-} = (S_{s} - K)^{-},$$
then we know that $(S_{t} - K)^{-}$ is a submartingale. Taking expectation on the both side, we have for all $s \leq t$,
$$\mathbb{E} [(S_{t} - K)^{-} ] \geq \mathbb{E} [(S_{s} - K)^{-}].$$
So, $P(t)$ is increasing.


\end{document}
